Optimal Duplicate Detection in a Sliding Window

### 2020-03-30 qhtv2.tex
Typos

	Building on the Bloom filter \cite{bloom_spacetime_1970}, an early data structure amenable to duplicate detection, several generalisations and variants were proposed tthat reduce error rates while maintaining constant memory.

	Unfortunately the infinite stream setting is ill-posed, and error rates of duplicate detection filters turn out to be heavily constrained: as a consequence they appear to provide no advantage, asymptotically, over a random biased coin toss \cite{10.1145/3297280.3297335}.

	Both as a better-behaved setting and as a construction technique, a
	\emph{windowed} variant of the duplicate detection problem was 
	introduced in \cite{She08,Yoo10}, which consists in finding duplicate 
	entries amongst the last $w$ elements of the stream for a fixed integer $w$.

	In this paper we formalise this setting, and show that a perfect (zero error) solution can be used up to a maximal window size $w_\text{max}$. Above this threshold we show that some existing duplicate detection filters (designed for the \emph{non-windowed} setting) perform better that those targeting the windowed problem. Finally, we introduce a \enquote{queuing construction} that improves on the performance of some duplicate detection filters in the windowed setting.

@@

\subsection{Contributions}
In this paper, we start from a naïve solution for the wDDP to then derive bounds for when it can be solved within $M$ bits of memory, up to a window size $w_\text{max}$, in constant time. We then introduce a generalisation of the naïve solution, and study its error rate. We show that this construction, which we call Short Hash Filter (SHF), can push the value $w_\text{max}$ further while operating at being constant time --at the cost of some errors. If allowing for linear time, we also provide a better alternative, the Compact Short Hash Filter (CSHF).

@@

Finally, we provide an analysis of our queueing construction's resistance to adversarial streams. (Section~\ref{sec:adversarial}).

\subsection{Related work}\label{sub:related}
The notion of sliding window was, as far as we know, first introduced in \cite{10.1145/1060745.1060753}. However, several variations exist and describe incompatible situations, such as \cite{shtul2020agepartitioned}. Instead, the wDDP formulation we rely on is due to \cite{Yoo10,She08}, which also introduce algorithms for solving the wDDP approximately.

The notion of using subfilters, as in the queuing construction, can be found in the A2 filter's design \cite{Yoo10} and a variation thereof can be found in \cite{shtul2020agepartitioned} but in a different DDP formulation. The A2 is built from two Bloom filters, a construction which we generalise and analyse generically in this paper. Similarly, the construction in \cite{shtul2020agepartitioned} only works with Bloom Filters.

@@

\begin{theorem}
	Let $E$ be a stream of $n$ elements uniformly selected from an alphabet of size $|\Gamma|$. For any DDF using $M$ bits of memory, the error probability $ER_n = \FP_n + \FN_n$ satisfies
	\begin{equation}
	E_n \geq 1 - \frac{1 - \left(1 - \frac{1}{|\Gamma|}\right)^{M}}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n} 
	\end{equation}
	for any $n > M$.
\end{theorem}
In particular, the asymptotic error rate $E_\infty$ satisfies $E_\infty \geq \left(1 - \frac{1}{|\Gamma|}\right)^{M} \approx 1 - M/|\Gamma|$.

@@

\section{Queuing filters from existing DDFs}
\label{sec:subfilter-select}
\label{sec:optimal_ddf}
Our queuing construction relies on a choice of subfilters. A first observation 
is that we may assume that all subfilters can be instances of a single 
DDF design (rather than a combination of different designs).

>>>>

	Building on the Bloom filter \cite{bloom_spacetime_1970}, an early data structure amenable to duplicate detection, several generalizations and variants were proposed that reduce error rates while maintaining constant memory.

	Unfortunately the infinite stream setting is ill-posed, and error rates of duplicate detection filters turn out to be heavily constrained: as a consequence they appear to provide no advantage, asymptotically, over a biased coin toss \cite{10.1145/3297280.3297335}.

	Both as a better-behaved setting and as a construction technique, a
	\emph{windowed} variant of the duplicate detection problem was 
	introduced in \cite{She08,Yoo10} which consists in finding duplicate 
	entries amongst the last $w$ elements of the stream for a fixed integer $w$.

	In this paper we formalize this setting, and show that a perfect (zero error) solution can be used up to a maximal window size $w_\text{max}$. Above this threshold we show that some existing duplicate detection filters (designed for the \emph{non-windowed} setting) perform better that those targeting the windowed problem. Finally, we introduce a \enquote{queuing construction} that improves on the performance of some duplicate detection filters in the windowed setting.

@@

\subsection{Contributions}
In this paper, we start from a naïve solution for the wDDP to then derive bounds for when it can be solved within $M$ bits of memory, up to a window size $w_\text{max}$, in constant time. We then introduce a generalization of the naïve solution, and study its error rate. We show that this construction, which we call Short Hash Filter (SHF), can push the value $w_\text{max}$ further while operating in constant time --- at the cost of some errors. We also provide a different tradeoff, the Compact Short Hash Filter (CSHF), which uses fewer memory but operates in linear time.

@@

Finally, we provide an analysis of our queueing construction's resistance to adversarial streams.

\subsection{Related work}\label{sub:related}
The notion of sliding window was, as far as we know, first introduced in \cite{10.1145/1060745.1060753}; but several variations exist that are incomparable to one another (e.g. \cite{shtul2020agepartitioned}). The wDDP formulation we rely on is due to \cite{Yoo10,She08}, which also introduce algorithms for solving the wDDP approximately.

The notion of using subfilters, as in the queuing construction, can be found in the A2 filter's design \cite{Yoo10} and a variation thereof can be found in \cite{shtul2020agepartitioned} but in a different DDP formulation. The A2 is built from two Bloom filters, a construction which we generalize and analyse generically in this paper. Similarly, the construction in \cite{shtul2020agepartitioned} only works with Bloom Filters.

@@

\begin{theorem}
	Let $E$ be a stream of $n$ elements uniformly selected from an alphabet of size $|\Gamma|$. For any DDF using $M$ bits of memory, the error probability $ER_n = \FP_n + \FN_n$ satisfies
	\begin{equation}
	E_n \geq 1 - \frac{1 - \left(1 - \frac{1}{|\Gamma|}\right)^{M}}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n} 
	\end{equation}
	for any $n > M$.
\end{theorem}
In particular, the asymptotic error rate $E_\infty$ satisfies 
\begin{equation*}
	E_\infty \geq \left(1 - \frac{1}{|\Gamma|}\right)^{M} \approx 1 - M/|\Gamma|.
\end{equation*}

@@

\subsection{Queuing filters from existing DDFs}
\label{sec:subfilter-select}
\label{sec:optimal_ddf}
Our queuing construction relies on a choice of subfilters. A first observation 
is that we may assume that all subfilters can be instances of a single 
DDF design (rather than a combination of different designs).
